---
title: "dchs2015"
author: "George Oberle"
date: "August 3, 2015"
output: html_document
---


#Institutionalizing Knowledge in Early Washington D.C.  

In Democracy in America, Alexander de Tocqueville observed that “In democratic countries the science of association is the mother of science; the progress of all the rest depends upon the progress it has made.” Tocqueville noted that Americans formed numerous associations for distinct purposes and whose goals included establishing schools and churches, and spreading knowledge through the publication of information of interest to like-minded citizens.  The residents of the District of Columbia, America’s federal city, sought to create a thriving metropolis crowded with a vibrant and extensive community of scholars. Many hoped a national institution would materialize to promote the expansion of knowledge and learning. This paper explores the purpose and composition of the myriad societies in the early federal city, especially scientific, literary and proto-professional societies, 1815-1850. This paper examines the shared relationships between these groups using computational tools and methods to develop network analysis using society membership lists.

##Introduction

In addition to being the site for the national government, the District of Columbia was the home of many different societies dedicated to the collection and dissemination of useful knowledge. Typically, scholars characterize these institutions as localized groups with grand designs limited by parochial insights and miserly pocketbooks. This paper seeks to explore the make-up of the local institutions and to utilize computational methods to examine if these local institutions were more cosmopolitan in nature. 

Introduction
Hook
Questions
What were the institutions and what did they do?
Who joined these institutions?
Were they interrelated?
Did they connect to organization outside of DC?

##Methods
  * Organizations
  * Membership Lists
	+ Columbian Agricultural Society
    + Columbian Institute for the Promotion of Arts and Sciences
    + DC Medical Society
    + National Institute of Science
  + Military Philosophical Society


###Membership Lists
<h3>Membership Rolls: Collecting Data</h3>
Finding and collecting data from the various organizations that existed in Early Washington is a tedious project and fraught with challenges. There are limited archival sources from the organizations themselves. Some of the more complete examples include the Columbian Institute for the Promotion of Arts and Sciences and its successor the National Institution (Institute) for the Promotion of Science both available in the Smithsonian Archives. These particular institutions have membership books that compiled the membership lists at different times in their existence. Since these institutions counted many leading citizens of the republic as members and they scholars identify them as predecessors of the Smithsonian many of their records survived. In other cases like the Medical Society of the District of Columbia’s records are at the National Library of Medicine. Occasionally organizations published the names of their membership in their journals or other proceedings such as the Columbian Agricultural Society.  Most typically, there seem to be hints or references to the associations in scattered newspaper accounts, memoirs, or government records. Despite the restrictions of the source pool, there are ways to utilize the existing data to ask new questions from old sources. This paper seeks to demonstrate the importance of such lists to scholars in order to help demonstrate how computational methods can provide new uses to old sources. Additionally, communicating these methods may help uncover more relevant source pools for scholars to mine.


Data

```{r}

#Load tools
library (sp)
library (rgdal)
library (rgeos)
library (maptools)
library (ggmap)
library (ggplot2)
library (mullenMisc)
library (magrittr)
library (stringr)
library (dplyr)
library (RColorBrewer)
library (igraph)
```

```{r}
#Load membership rollls data

data = read.csv("~/Desktop/DCHS2015/data/learnsoc1.csv")

datanew = read.csv("~/Desktop/DCHS2015/data/newdecnetwork.csv")

```

After reading the data this sample set of membership lists from a variety of different learned societies between x and y. The list has 2262 members lists coming from x organizations.

```{r}
data_members <- graph.data.frame(data, directed = FALSE)
plot (data_members, vertex.label = NA)
title("Test", sub = "A network graph of members",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 0.75, font.sub = 3, col.sub = "red")

dc_data_members <- graph.data.frame(datanew, directed = FALSE)
plot (dc_data_members, vertex.label = NA)
title("Test", sub = "A network graph of members",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 0.75, font.sub = 3, col.sub = "red")

```


```{r}
data_members <- graph.data.frame(data, directed = FALSE)
plot (data_members)
title("Test", sub = "A network graph of members with name labels",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 0.75, font.sub = 3, col.sub = "red")


```

```{r}
V(data_members)
bipartite.mapping(data_members)
V(data_members)$type <- bipartite.mapping(data_members)$type
is.bipartite(data_members)
plot(data_members, layout = layout.bipartite, vertex.label = NA)
title("Test", sub = "A bipartite layout of members and organizations",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 0.75, font.sub = 3, col.sub = "red")

#V(dc_data_members)
#bipartite.mapping(dc_data_members)

V(dc_data_members)$type <- bipartite.mapping(dc_data_members)$type
is.bipartite(dc_data_members)
plot(dc_data_members, layout = layout.bipartite, vertex.label = NA)
title("Test", sub = "A bipartite layout of members and organizations",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 0.75, font.sub = 3, col.sub = "red")


```


```{r}
full_members <- bipartite.projection(data_members)

full_members_DC <- bipartite.projection(dc_data_members)

```


```{r}
#Just the organizations
plot(full_members$proj2)
title("Test", sub = "Just the Organizations",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 2, font.sub = 3, col.sub = "red")

plot(full_members_DC$proj2)
title("Test", sub = "Just the Organizations",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 2, font.sub = 3, col.sub = "red")
```

```{r}
E(full_members$proj2)
E(full_members$proj2)$weight
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
range01(E(full_members$proj2)$weight)

E(full_members_DC$proj2)
E(full_members_DC$proj2)$weight
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
range01(E(full_members_DC$proj2)$weight)
```

This Range makes the scale smaller to improve viewing.
```{r}
plot(full_members$proj2, edge.width = range01(E(full_members$proj2)$weight) * 20)
title("Organizations", sub = "Weighted by Shared Members",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 2, font.sub = 3, col.sub = "red")

plot(full_members_DC$proj2, vertex.size = 10, layout= layout.fruchterman.reingold,edge.curved=TRUE,vertex.color="blue", edge.color="red", edge.width = range01(E(full_members_DC$proj2)$weight) * 20)
title("Organizations", sub = "Weighted by Shared Members",cex.main = 2,   font.main= 4, col.main= "blue", cex.sub = 2, font.sub = 3, col.sub = "red")

```

```{r}
#Count

datanew %>%
  group_by(Organization) %>%
  summarize(n = n()) %>%
  filter(n > 1) %>%
  arrange(desc(n))

```


```{r}
#Load Silliman's Journal data
datascisoc_1825 = read.csv("~/Desktop/DCHS2015/data/Scientific_Societies_1825_American_Journal_of_Science_and_Arts.csv")
```

```{r}
#series of functions that add new variable called location by combining City and State and then use the geocode api from google. Then bind the two dataframes together and save them.
datascisoc_1825<- datascisoc_1825%>%
mutate(location=paste (City, State))
location_geocoded <- geocode(datascisoc_1825$location, source = "google")
datascisoc_1825_geocoded <- cbind(datascisoc_1825, location_geocoded)
datascisoc_1825_geocoded
write.csv(datascisoc_1825_geocoded, file = "~/Desktop/DCHS2015/data/datascisoc_1825_geocoded.csv")

```



```{r}
#This is a bar graph of the number of societies established by date reported by Silliman.
state <- ggplot (datascisoc_1825_geocoded, aes (x=State))
state + geom_bar (width=.5, fill="blue") + xlab ("State")

year <- ggplot (datascisoc_1825_geocoded, aes (x=Year.Established)) 
year + geom_bar (fill="red") + xlab ("Year Established")) +ggtitle ("Learned Societies Reported in Silliman's Journal-1825")
#To do Need to figure out better x scale label to show all dates.


org <- ggplot(data, aes(Organization))
org + geom_bar(width=.5, fill="blue", color="darkblue") + coord_flip() +ggtitle ("Total Members of Societies DC relations")

dcorg <- ggplot (datanew, aes(Organization), scale_x_log10())
dcorg + geom_bar(width=.5, fill="blue", color="darkblue") + coord_flip() +ggtitle ("Total Members of Societies DC relations")
#log scale not working

```


##Appendix 1 List of Societies in Washington D.C.
Columbian Institute
Washington D.C. 


Methods
  Organizations
	Membership Lists
		Columbian Agricultural Society
		Columbian Institute for the Promotion of Arts and Sciences
		National Institute of Science
	
Analysis

Conclusion



Columbian Agricultural Society
Columbian Institute for the Promotion of Arts and Sciences
Medical Society of the District of Columbia
Enosinian Society –Columbian College
Washington Relief Society 1830
Female Union Benevolent Society of Washington City
Washington Monument Society
Washington Literary Society
Washington Library Company
Washington Female Orphan Society
Washington Orphan Asylum
Washington Botanical Society
Washington Bible Society
Washington Benevolent Society of Young Men
Washington Art League
Washington Art Association
Columbia Typographical Society
Odd Fellows
Grand United Order of Odd fellows
Grand United Order of colored Odd fellows
National Institution (Institute) for the Promotion of Science
Mechanics Institute
Consumers Protective Association
Columbian Institution for the Deaf, Dumb and Blind
Columbian Hospital for Women and Lying in Asylum
American Tract Society
American Colonization Society
  The establishment of the seat of government 
nation’s capital in Washington D.C. 

Data
Columbian Institute
Washington D.C. 
Introduction
Hook
Questions
What were the institutions and what did they do?
Who joined these institutions?
Were they interrelated?
Did they connect to organization outside of DC?

Methods
	Organizations
	Membership Lists
		Columbian Agricultural Society
		Columbian Institute for the Promotion of Arts and Sciences
		National Institute of Science
	
Analysis

Conclusion






```{r}
#I would like to use this to compare with development of other institutions. 
milit_soc <- read.csv("~/Desktop/DCHS2015/data/militacadlist.csv")
milit_soc<- milit_soc%>%
mutate(location=paste (City, State))
location_geocoded <- geocode(milit_soc$location)
milit_soc_geocoded <- cbind(milit_soc, location_geocoded)
milit_soc_geocoded
write.csv(milit_soc_geocoded, file = "milit_soc_geocoded.csv")

```

```{r}
#Load data
schol = read.csv("~/Desktop/DCHS2015/data/schol-soc-geocoded.csv")
coll = read.csv("~/Desktop/DCHS2015/data/early_colleges_geocoded.csv")
militgeo = read.csv("~/Desktop/DCHS2015/data/milit_soc_geocoded.csv")
#Preparing 1830 Map
map_sp <- readOGR("nhgis-shp/", "state_1830")
class(map_sp)
str(map_sp, max.level = 2)
map_sp@bbox
map_sp@proj4string
head(map_sp@data)
plot(map_sp)

map_sp1 <- readOGR("nhgis-shp/", "state_1840")
class(map_sp1)
str(map_sp1, max.level = 2)
map_sp1@bbox
map_sp1@proj4string
head(map_sp1@data)
plot(map_sp1)

map1830<- readOGR ("nhgis-shp" , "state_1830")
map1840<- readOGR ("nhgis-shp" , "state_1840")
map1850<- readOGR ("nhgis-shp" , "state_1850")

fort<- fortify (map1830)
fort2<- fortify (map1840)
fort3<- fortify (map1850)

ggplot () + geom_map(data=fort, map=fort, aes(x=long, y=lat, group=group, map_id=id), color="black", fill="white") +geom_point(data=coll, aes (x=lon, y=lat), color="red") + coord_map()

ggplot () + geom_map(data=fort3, map=fort3, aes(x=long, y=lat, group=group, map_id=id), color="black", fill="white") +geom_point(data=militgeo, aes (x=lon, y=lat), color="red") + coord_map()

ggplot () + geom_map(data=fort, map=fort, aes(x=long, y=lat, group=group, map_id=id), color="black", fill="white") +geom_point(data=coll, aes (x=lon, y=lat, color=factor(established)) ) + coord_map()

ggplot () + geom_map(data=fort, map=fort, aes(x=long, y=lat, group=group, map_id=id), color="black", fill="white") +geom_point(data=coll, aes (x=lon, y=lat, color=factor(period), shape=factor(period)) ) + coord_map()

ggplot () + geom_map(data=fort2, map=fort2, aes(x=long, y=lat, group=group, map_id=id), color="black", fill="white") +geom_point(data=coll, aes (x=lon, y=lat, color=factor(period), shape=factor(period)) ) + coord_map()
 
```

```{r}
library(leaflet)

leaflet() %>%
  addTiles() %>%
  addPolygons(data = map1850, fillOpacity = 0, weight = 1, color = "black") %>%
  addCircleMarkers(data = schol, color = "blue", fillOpacity = 0) %>%
  addCircleMarkers(data = coll, color = "red") %>%
  addCircleMarkers(data = militgeo, color = "green")
```



```{r}
libraries <- read.csv("~/Desktop/DCHS2015/data/libraries_davies.csv")
periodize <- function(founding_date) {
  period <- NA
  period <- ifelse(founding_date <= 1776, "Colonial", period)
  period <- ifelse(founding_date >= 1776 & founding_date <= 1828, "Early National", period)
  period <- ifelse(founding_date >1828, "Antebellum", period)
  return(period)
  period <- ifelse(founding_date >1860, "War and Postwar", period)
  return(period)
  }
  
libraries <- libraries %>%
mutate(period = periodize(founding_date))
write.csv(libraries, file = "libraries_period.csv")

libraries_period <- read.csv("~/Desktop/DCHS2015/libraries_period.csv")
#This loads a list of 2000 Libraries from the list of 8528 gathered from the Davies Project. The original had 16 variables. This process combines the locality and state into one variable making lib_per1 into 17 variables. Finally libper1_geocoded uses the Google api to geocode the items as a process yeilding 2 new variables. Using the same named function bind the lib_per1 and libper1_geocoded into one file and save it to the drive.
lib_per1 <- read.csv("~/Desktop/DCHS2015/libperiod1.csv")
lib_per1<- lib_per1%>%
mutate(location=paste (locality, state))
libper1_geocoded <- geocode(lib_per1$location, source = "google")
libper1_geocoded <- cbind(lib_per1, libper1_geocoded)
libper1_geocoded
write.csv(libper1_geocoded, file = "libper1_geocoded.csv")






milit_soc <- read.csv("~/Desktop/DCHS2015/data/militacadlist.csv")
milit_soc<- milit_soc%>%
mutate(location=paste (City, State))
location_geocoded <- geocode(milit_soc$location)
milit_soc_geocoded <- cbind(milit_soc, location_geocoded)
milit_soc_geocoded
write.csv(milit_soc_geocoded, file = "milit_soc_geocoded.csv")
State Born for the Government Data.


Mapping The Larger Picture. Need Library Dataset and add above military.




#create period function
periodize <- function(year) {
  period <- NA
  period <- ifelse(year >= 1776, "Colonial", period)
  period <- ifelse(year >= 1776 & year <= 1828, "Early National", period)
  period <- ifelse(year >1828, "Antebellum", period)
  return(period)
  
 #These are the processes to use to add columns for geocoded.
early_colleges<- early_colleges%>%
mutate(location=paste (city, state))
location_geocoded <- geocode(early_colleges$location)
early_colleges_geocoded <- cbind(early_colleges, location_geocoded)
early_colleges_geocoded


mutate(location=paste (City, State))
location_geocoded <- geocode(milit_soc$location)
milit_soc_geocoded <- cbind(milit_soc, location_geocoded)
milit_soc_geocoded
write.csv(milit_soc_geocoded, file = "milit_soc_geocoded.csv")
  early_colleges_geocoded %>%
 # mutate(period = periodize(established))
#Thanks to Colleagues